---
#Reproducible Research: Assignment 1


### Loading and processing the data
```{r}
datos <- read.csv("./repdata_data_activity/activity.csv")
no_NA <- na.omit(datos)
rownames(no_NA) <- 1:nrow(no_NA)
head(no_NA)
```
###Calculation of the mean and median total number of steps taken per day 

Mean total number of steps taken per day
```{r}
total_number_steps <- aggregate(no_NA$steps, list(Date = no_NA$date), FUN = "sum")$x
mean(total_number_steps)
```
Median total number of steps taken per day
```{r}
median(total_number_steps)
```
###Histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(total_number_steps, xlab='Total steps', ylab='Frequency')
```

###Average daily activity pattern

Calculation of  the mean steps for each five minute interval, and then put it in a data frame.

```{r}
steps_interval <- tapply(datos$steps, INDEX = datos$interval, FUN = mean, na.rm = TRUE)
head(steps_interval)
hour<-1:length(steps_interval)/12
layout(mat = 1);
```

```{r}
qplot(y=steps_interval, x=hour, geom="line", xlab="5 Min Interval", ylab="Number of steps", main="Average number of steps taken along the day")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max(steps_interval)
steps_interval[100:110]
```
The 5-minute interval that, on average across all the days in the dataset, contains the maximum number of steps is 835.

###Imputing missing values 

The missing values are indicated with a NA in the original dataset. The total number of NA values s given by: 

```{r}
sum(is.na(datos$steps));
```
The percentage of missing values is:
```{r}
100*sum(is.na(datos$steps)) / length(datos$steps);
```
##Devise a strategy for filling the missing values

A very simple method is to take the mean of a 5 min interval thru all days to fill ne NA values in the corresponding interval

```{r}
datos$fill <- datos$steps;
filling <- rep(steps_interval,
               times = length(datos$steps) / length(steps_interval));
for (i in 1:length(datos$steps)) {
    if (is.na(datos$fill[i]))
        datos$fill[i] <- filling[i];
    }
```
We proceed to do the same calculations with the filled data.  A comparison between the two summaries (with and without filling the NA values)

```{r}
steps_date_fill <- tapply(datos$fill, INDEX = datos$date,
                             FUN = sum);
summary(steps_date_fill);
summary(total_number_steps);
```
shows that both results have the same mean and median. The histogram of the filled data shows that the original data distribution possesses fatter tails and the maximum is smaller.  

```{r}
qplot(steps_date_fill, xlab='Total steps', ylab='Frequency');
```
